---
title: 'Class 17: Covid-19 Vaccination Rates'
author: "Dominique Lie (A15470100)"
date: "11/23/2021"
output:
  pdf_document: default
  html_document: default
---

Background 

The goal of this hand-on mini-project is to examine and compare the Covid-19 vaccination rates around San Diego. 

```{r}
# Import vaccination data
vax <- read.csv("covid19.csv")
head(vax)
```
## Ensure the data column is useful

We will use the **lubridate** package to make life a lot easier when dealing with dates and times

```{r}
library(lubridate)
```

```{r}
today()
```

Here we make our 'as_of_date' column lubridate format...

```{r}
#specify that we are using the Year-month-day format
vax$as_of_date <- ymd(vax$as_of_date)
```

```{r}
today() - vax$as_of_date[1]
```


> Q1. What column details the total number of people fully vaccinated? 

```{r}
colnames(vax)
```
column 9

> Q2. What column details the Zip code tabulation area?

column 2

> Q3. What is the earliest date in this dataset?

```{r}
vax$as_of_date[1]
```
earliest is 01/05/2021

> Q4. What is the latest date in this dataset? 

```{r}
nrow(vax)
vax$as_of_date[81144]
```

lastest is 11/16/2021

# Skimr

As done previously, call skim() function from skimr package to get quick overview of data. 

```{r}
skimr::skim(vax)
```
> Q5. How many numeric columns are in this dataset? 

There are 9 numeric columns

> Q6. Note that there are "missing values" in the dataset. How many NA values are there in the persons_fully_vaccinated columns? 

```{r}
sum(is.na(vax$persons_fully_vaccinated))
```

There are 8256 missing values

> Q7. What percent of persons_fully_vaccinated values are missing (to 2 significant figures)?

```{r}
signif((sum(is.na(vax$persons_fully_vaccinated)))/nrow(vax), 2)
```


> Q8. Why might this data be missing?

The data could be redacted for privacy 


> Q9. How many days since the first entry and the last entry?

```{r}
vax$as_of_date[nrow(vax)] - vax$as_of_date[1]
```

```{r}
today() - vax$as_of_date[nrow(vax)]
```

> Q10. How many unique dates are in the dataset (i.e. how many different dates are detailed)? 

```{r}
length(unique(vax$as_of_date))
```

This sounds good 

```{r}
46 * 7
```

## Working with ZIP codes

```{r}
library(zipcodeR)
```

```{r}
geocode_zip("92037")
```
```{r}
zip_distance('92037', '92109')
```
```{r}
reverse_zipcode(c('92037', '92109'))
```
```{r}
# Pull data for all ZIP codes in the dataset
zipdata <- reverse_zipcode( vax$zip_code_tabulation_area )
```


# Focus on SD county

```{r}
table(vax$county)
```
We will subset with base R

```{r}
inds <- vax$county == "San Diego"

head(vax[inds, ])
```

```{r}
library(dplyr)
```
```{r}
sd <- filter(vax, county == "San Diego")
nrow(sd)
```

```{r}
sd.10 <- filter(vax, county == "San Diego" &
                age5_plus_population > 10000)
```

> Q11. How many distinct zip codes are listed for San Diego County? 

```{r}
length((unique(sd$zip_code_tabulation_area)))
```

> Q12. What San Diego County Zip Code has the largest 12+ Population in this dataset? 

```{r}
which.max(sd$age12_plus_population)
sd[23, ]

#or

inds <- which.max(sd$age12_plus_population)
sd[inds,]
```
92154

What is the population in the 92037 ZIP code area? 

```{r}
filter(sd, zip_code_tabulation_area == 92037)[1, ]
```
> Q13. What is the overall average “Percent of Population Fully Vaccinated” value for all San Diego “County” as of “2021-11-09”?

```{r}
sd.now <- filter(vax, county == "San Diego", as_of_date == "2021-11-09")

mean(sd.now$percent_of_population_fully_vaccinated, na.rm = TRUE)
```

We can look at the 6-number summary 

```{r}
summary(sd.now$percent_of_population_fully_vaccinated)
```


> Q14. Using either ggplot or base R graphics make a summary figure that shows the distribution of Percent of Population Fully Vaccinated values as of “2021-11-09”?

Using base R plot

```{r}
hist(sd.now$percent_of_population_fully_vaccinated)
```

Using ggplot

```{r}
library(ggplot2)
```

```{r}
ggplot(sd.now) + aes(percent_of_population_fully_vaccinated) + geom_histogram(bins = 15)
```

What about 92037 - UCSD/La Jolla? 

```{r}
ucsd <- filter(sd, zip_code_tabulation_area == "92037")
ucsd[1,]$age5_plus_population
```

> Q15. Using ggplot make a graph of the vaccination rate time course for the 92037 ZIP code area: 

```{r}
ggplot(ucsd) +
  aes(as_of_date,
      percent_of_population_fully_vaccinated) +
  geom_point() +
  geom_line(group=1) +
  ylim(c(0,1)) +
  labs(x = "Date", y="Percent Vaccinated")
```

## Comparing 92037 to other similar sized areas? 

```{r}
#Subset to all CA areas with a population as large as 92037 
vax.36 <- filter(vax, age5_plus_population > 36144 &
                as_of_date == "2021-11-16")
head(vax.36)
```
> Q16. Calculate the mean "Percent of Population Fully vaccinated" for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date "2021-11-16". Add this as a striaght horizontal line to your plot from above with the geom_hline() function

```{r}
int <-mean(vax.36$percent_of_population_fully_vaccinated)
```

```{r}
p <- ggplot(ucsd) +
  aes(as_of_date, percent_of_population_fully_vaccinated) +
  geom_point() +
  geom_line(group=1) +
  ylim(c(0,1)) +
  labs(x = "Date", y="Percent Vaccinated")
p + geom_hline(yintercept = int, linetype = "dashed")
```

> Q17. What is the 6 number summary (Min, 1st Qu., Median, Mean, 3rd Qu., and Max) of the “Percent of Population Fully Vaccinated” values for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2021-11-16”?

```{r}
summary(vax.36$percent_of_population_fully_vaccinated)
```

> Q18. Using ggplot generate a histogram of this data.

```{r}
ggplot(vax.36) + aes(percent_of_population_fully_vaccinated) + geom_histogram()
```

> Q19. Is the 92109 and 92040 ZIP code areas above or below the average value you calculated for all these above?

```{r}
vax %>% filter(as_of_date == "2021-11-16") %>%  
  filter(zip_code_tabulation_area=="92040") %>%
  select(percent_of_population_fully_vaccinated)
```
92040 is below average

```{r}
vax %>% filter(as_of_date == "2021-11-16") %>%  
  filter(zip_code_tabulation_area=="92109") %>%
  select(percent_of_population_fully_vaccinated)
```
92109 is above average

> Q20. Finally make a time course plot of vaccination progress for all areas in the full dataset with a age5_plus_population > 36144.

```{r}
vax.36.all <- filter(vax, age5_plus_population > 36144)


ggplot(vax.36.all) +
  aes(as_of_date,
      percent_of_population_fully_vaccinated, 
      group=zip_code_tabulation_area) +
  geom_line(alpha=0.2, color="blue") +
  labs(x= "Date", y= "Percent Vaccinated",
       title = "Vaccination rate across California",
       subtitle = "Only areas with population above 36k are shown") +
  geom_hline(yintercept = 0.66, linetype = "dashed")
```

> Q21. How do you feel about traveling for Thanksgiving and meeting for in-person class next Week?

okay for remote Tuesday and in-person the rest of the week
