---
title: "class10"
author: "Dominique Lie (A15470100)"
date: "10/28/2021"
output:
  pdf_document: default
---
# 1. Importing candy data
```{r}
candy_file <- "https://raw.githubusercontent.com/fivethirtyeight/data/master/candy-power-ranking/candy-data.csv"
candy = read.csv(candy_file, row.names = 1)
head(candy)
```
> Q1. How many different candy types in this dataset?

```{r}
nrow(candy)
```

> Q2. How many fruity candy types in the dataset?

```{r}
sum(candy$fruity)
```

# 2. What is your favorite candy? 

```{r}
candy["Twix", ]$winpercent
```
> Q3. What is your favorite candy in the dataset and what is it's winpercent value? 

```{r}
candy["Red vines", ]$winpercent
```

> Q4. What is the winpercent value for "Kit Kat"?

```{r}
candy["Kit Kat", ]$winpercent
```

> Q5. What is the winpercent value for "Tootsie Roll Snack Bars"?

```{r}
candy["Tootsie Roll Snack Bars", ]$winpercent
```

```{r}
# install.packages("skimr")
library("skimr")
skim(candy)
```
> Q6. Is there any variable/column that looks to be on a different scale to the majority of the other columns in the dataset?

winpercent - on percentage scale to 100 rather than to 1

> Q7. What do you think a zero and one represent for the candy$chocolate column?

A zero means that it is not chocolate and one means that it is chocolate. 

> Q8. Plot histogram of winpercent values

```{r}
hist(candy$winpercent)
```
> Q9. Is the distribution of winpercent values symmetrical?

No, it is slightly skewed to the left. 

> Q10. Is the center of the distribution above or below 50%? 

The distribution is below 50%. 

> Q11. On average is chocolate candy higher or lower ranked than fruit candy? 

```{r}
choc <- candy$winpercent[as.logical(candy$chocolate)]
mean(choc)
fruit <- candy$winpercent[as.logical(candy$fruity)]
mean(fruit)
```

On average, the chocolate candy is higher ranked than fruit candy.

> Q12. Is this different statistically significant?

```{r}
t.test(choc, fruit)
```
The difference is signficantly significant as the p-value is less than 0.05.

# 3. Overall Candy Rankings

> Q13. What are the five least liked candy types in this set?

```{r}
head(candy[order(candy$winpercent),], n = 5)
```

> Q14. What are the top 5 all time favorite candy types out of this set?

```{r}
head(candy[order(candy$winpercent, decreasing = TRUE), ], n = 5)
```

> Q15. Make a first barplot of candy ranking based on winpercent values. 

```{r}
library(ggplot2)
ggplot(candy) + aes(winpercent, rownames(candy)) + geom_bar(stat = "identity")
```

> Q16. This is quite ugly, use the reorder() function to get the bars sorted by winpercent? 

```{r}
ggplot(candy) + aes(winpercent, reorder(rownames(candy), winpercent)) + geom_bar(stat = "identity")
```
Set up color

```{r}
my_cols=rep("black", nrow(candy))
my_cols[as.logical(candy$chocolate)] = "chocolate"
my_cols[as.logical(candy$bar)] = "brown"
my_cols[as.logical(candy$fruity)] = "pink"

```

```{r}
ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col(fill=my_cols) 
```
# Taking a look at pricepercent

```{r}
library(ggrepel)

# How about a plot of price vs win
ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=my_cols) + 
  geom_text_repel(col=my_cols, size=3.3, max.overlaps = 5)
```

> Q19. Which candy type is the highest ranked in terms of winpercent for the least money - i.e. offers the most bang for your buck?

Reese's O Miniatures


> Q20. What are the top 5 most expensive candy types in the dataset and of these which is the least popular?

```{r}
ord <- order(candy$pricepercent, decreasing = TRUE)
head( candy[ord,c(11,12)], n=5 )
```

> Q21. Make a barplot again with geom_col() this time using pricepercent and then improve this step by step, first ordering the x-axis by value and finally making a so called “dot chat” or “lollipop” chart by swapping geom_col() for geom_point() + geom_segment().

```{r}
ggplot(candy) + aes(pricepercent, reorder(rownames(candy), pricepercent)) + geom_col()
```
```{r}
ggplot(candy) +
  aes(pricepercent, reorder(rownames(candy), pricepercent)) +
  geom_segment(aes(yend = reorder(rownames(candy), pricepercent), 
                   xend = 0), col="gray40") +
    geom_point()
```
# 5. Exploring correlation structure

```{r}
library(corrplot)
cij <- cor(candy)
corrplot(cij)
```
> Q22. Examining this plot what two variables are anti-correlated (i.e. have minus values)?

fruity and chocolate
> Q23. Similarly, what two variables are most positively correlated?

winpercent and chocolate

# 6 Principal Component Analysis

```{r}
pca <- prcomp(candy, scale = TRUE)
summary(pca)
```
```{r}
plot(pca$x[,1:2])
```
```{r}
plot(pca$x[,1:2], col=my_cols, pch=16)
```
```{r}
# Make a new data-frame with our PCA results and candy data
my_data <- cbind(candy, pca$x[,1:3])

p <- ggplot(my_data) + 
        aes(x=PC1, y=PC2, 
            size=winpercent/100,  
            text=rownames(my_data),
            label=rownames(my_data)) +
        geom_point(col=my_cols)

p


```

```{r}
library(ggrepel)

p + geom_text_repel(size=3.3, col=my_cols, max.overlaps = 7)  + 
  theme(legend.position = "none") +
  labs(title="Halloween Candy PCA Space",
       subtitle="Colored by type: chocolate bar (dark brown), chocolate other (light brown), fruity (red), other (black)",
       caption="Data from 538")
```

```{r}
library(plotly)
```
```{r}
ggplotly(p)
```
```{r}
par(mar=c(8,4,2,2))
barplot(pca$rotation[,1], las=2, ylab="PC1 Contribution")
```
> Q24. What original variables are picked up strongly by PC1 in the positive direction? Do these make sense to you?

fruity, hard, and pluribus
yes,it makes sense because a lot of candies that are fruity are hard and come in a package of more than one
